Learning by Teaching a Robot:\\The Case of Handwriting

### 2015-12-11 main.tex
Expanded discussion on ethical concerns

The place and role of the robot \textit{vis-\`a-vis} the teacher can be
questioned: as we see it, the role of the robot within the classroom (or at the
therapist's surgery) does not infringe upon the role of the adult (teacher or
therapist).  The core of the \emph{learning by teaching} paradigm relies on the
child becoming the teacher of an underperforming pupil (the robot): from that
perspective, the robot does not replace the teacher, on the contrary. It plays a
different role in the classroom, which happen to be novel as well: the robot is
the \emph{least} performing student, and still a very patient, always eager to
improve, one.

@@

Beyond handwriting, we however do believe that this work provides a novel
perspective on the role for robots in the field of education. \emph{Learning by
teaching} is a powerful paradigm because of not only its pedagogical efficacy,
but its potential to positively impact the child's motivation and self-esteem.
We hope that this article shows that this is a very relevant context of use for
robots: when facing a child with school difficulties, robots can play the role
of a na\"ive learner which neither adults nor peers -- because of the social
effects it would induce -- can convincingly play.

>>>>

\paragraph{Possible Ethical Concerns} Two aspects of this research ought to be
discussed in term of their possible ethical implications: the perceived role of
the robot \textit{vis-à-vis} the teacher, and the implications of the
mentor-protégé relationship for children, especially vulnerable ones.

The place and role of the robot \textit{vis-à-vis} the teacher can be
questioned: as we see it, the role of the robot within the classroom (or at the
therapist's surgery) does not infringe upon the role of the adult (teacher or
therapist).  The core of the \emph{learning by teaching} paradigm relies on the
child becoming the teacher of an underperforming pupil (the robot): from that
perspective, the robot does not replace the teacher, on the contrary. It plays a
different role in the classroom, which happen to be novel as well: the robot is
the \emph{least} performing student, and still a very patient, always eager to
improve, one.

@@

The implication of the mentor-protégé relationship on the children is less
clearly understood. The children can establish strong affective bonds with the robot (as
witnessed by the letter sent by Vincent several months after he interacted with
the robot), but we are not yet able to precisely characterize these bonds. The
ethical implication of the mentor-protégé relationship have been explored before
in the context of human teaching~\cite{brad1999mentor,wendelyn2008context}, but
they mostly look at the question from the perspective of the protégé, whereas in our
case, the child is the mentor. As such, relatively little is known on the
psychological implications for a child to commit in helping a robot, and as
advocated by Belpaeme and Morse~\cite{belpaeme2010time}, we likely need to first
gain more field experience before being able to draw conclusions.

Beyond handwriting, we however do believe that this work provides a novel
perspective on the role for robots in the field of education. \emph{Learning by
teaching} is a powerful paradigm because of not only its pedagogical efficacy,
but its potential to positively impact the child's motivation and self-esteem.
We hope that this article shows that this is a very relevant context of use for
robots: when facing a child with school difficulties, robots can play the role
of a naïve learner which neither adults nor peers -- because of the social
effects it would induce -- can convincingly play.
